\documentclass{article}
\usepackage{fancyhdr}
\usepackage{url}
\pagestyle{fancy}

\lhead{}
%\chead{Solving Games Symbolically}
\rhead{\thepage}
\lfoot{}
\cfoot{Introduction to {\sc Tailor}}
\rfoot{}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

%%\thispagestyle{plain}

%%\setlength{\oddsidemargin}{0.3in}
%%\setlength{\evensidemargin}{0.3in}

\begin{document}

\title{{\sc Tailor} Tutorial}

\author{Andrea Rendl}


\date{\today}

\maketitle
%%\begin{center}

\section{Introduction}
This tutorial introduces the tool {\sc Tailor} that 
converts constraint problem models formulated in the solver-independent modelling language 
{\sc Essence}$'$ to the input format of the Constraint Solver
{\sc Minion}. 
%\subsection{Modelling in Constraint Programming}
%To solve a problem with Constraint Programming techniques, 
%the problem has to be formulated as a Constraint Satisfaction 
%Problem(CSP) [***]. A  CSP consists of a set of variables
%and a set of constraints (relations) on these variables.
%Each variable ranges over a finite integer domain which the
%constraints restrict. A constraint can be any kind of 
%relation, using a wide range of operators (linear, non-linear,
%arithmetic, logic, ...). To describe CSPs more generally, 

{\sc Essence}$'$ is a solver-independent modelling 
language for Constraint Programming. It provides means to define 
variables, constants, parameters and offers a great range of 
constraint expressions, including complex constructs, 
such as quantifications.

{\sc Minion} is a fast scalable Constraint solver. 
However, modelling problems in {\sc Minion}'s input
language is time-consuming and tedious because of its 
primitive structure (it can be compared to writing a 
complex program in machine language).

{\sc Tailor} converts {\sc Essence}$'$ problem instances
into {\sc Minion} format and applies reformulatios  
(such as common subexpression elimination) during this 
process to enhance the problem model.

\subsection{Installing {\sc Tailor}}
{\sc Tailor} comes together with {\sc Minion} on
\url{http://minion.sourceforge.net/} as an executable 
Java {\tt jar}~\cite{java} file. It can also be downloaded
as a standalone version from 
\url{http://www.cs.st-and.ac.uk/~andrea/tailor/tailor.tar.gz}.

\subsection{Running {\sc Tailor}}
The Java jar file {\tt tailor.jar} can be executed either 
by double-clicking or in the command line with the command
\begin{center}
{\tt java -jar tailor.jar}
\end{center}
This initiates the graphical user interface of {\sc Tailor}. 
More details about the command line version are given with 
\begin{center}
{\tt java -jar tailor.jar -help}
\end{center}

\section{An Introduction to {\sc Essence}$'$}

%It allows you to formulate problem classes
%and problem instances. 

\subsection{Types and Domains}

Types and domains play a similar role; they prescribe a range of
values that a variable can take.
Types denote non-empty
sets that contain all elements
that have a similar structure, 
whereas domains denote possibly  empty
sets drawn from a
single type.  In this manner, each domain is associated with an
underlying type.  For example integer is the type underlying the
domain comprising integers between 1 and 10.

% what kind of types: only integers, booleans 
{\sc Essence}$'$ is a strongly typed language;
every expression
has a type, and the types of all
expressions can be inferred and checked for correctness.
Furthermore, {\sc Essence}$'$ is a finite-domain language; every decision variable is
associated with a finite domain of values. 


The 
atomic types of {\sc Essence}$'$ are {\tt int} (integer), {\tt bool} (Boolean) and
 user-defined enumerated types. There is also a compound type, array (or matrix),
 type that is constructed of atomic types.


There are three different types of domains in {\sc Essence}$'$:
boolean, integer and matrix/array domains. Boolean and integer 
domains are both atomic domains; array domains are built from 
atomic domains.

 Decision variables and quantified
variables need to be associated with a {\em finite} domain. 
The infinite integer domain, {\tt int}, is only valid with parameters. 
Decision variables are not allowed as domain elements.

\begin{description}

\item[Boolean Domains]
 {\tt bool} is the
 Boolean domain consisting of {\em false} and 
{\em true}. 

\item[Integer Domains]
An integer domain is a 
range of integers that can be either continuous,  e.g. $int$(1..10), or
sparse, e.g. $int$(1,3,5). Contineous domains are considered to be the empty 
domain if the lower bound is greater than the upper bound, such as in 
$int$(10..1).
The elements of sparse domains need to be ordered, hence $int$(1,5,3) 
is not valid. %and might produce error messages.


\item[Array Domains] 
%Arrays are used to combine a set of variables or
  %constant values. In {\sc Essence}$'$ 
  An array is defined by 
  the keyword {\tt matrix}, followed by its
  dimension and the base domain (over which the variables range).
  For instance,
  \begin{center}
   $NAME_1$ : {\tt matrix indexed by } [$int$(1..10)] {\tt of} $int$(1..5) 
  \end{center}
 stands for a 1-dimensional array of 10 elements where each element
 ranges from  $int$(1..5). The index domain states how to dereference 
 arrays. In the example above, $NAME_1$[1] dereferences the first
 element because the index domain starts with 1. 
 However, consider $NAME_2$ with a different index 
 domain, $int$(0..9):
  \begin{center}
   $NAME_2$ : {\tt matrix indexed by } [$int$(0..9)] {\tt of} $int$(1..5) 
  \end{center}
   $NAME_2$ is also a 1-dimensional array with 10 Boolean elements,
  but is dereferenced differently: $NAME_2$[1] dereferences the second 
  element because its index domain starts with 0.
  


%\item[Parameters]

%\item[Constants]

%\item[Variables]


\end{description}



\subsection{Basic Model Structure}
An {\sc Essence}$'$ model is structured in the following way:

\begin{enumerate}
\item Header with version number: {\tt  language Essence' 1.b.a} 
\item Parameter declarations (optional)
\item Constant definitions (optional)
\item Variable declarations (optional)
\item Objective (optional)
\item Constraints (optional)
\end{enumerate}

The version number needs a little explanation.   Over time it is inevitable that the input language will change, and the basic reason for the version number is to ensure that the user is protected against changes to the format which might change the semantics.   At all times Tailor should deal with a given version as intended, even if the latest version of the input format would be treated differently.   In the worst case, a future version of Tailor may refuse to process an earlier language format if the changes are too drastic.\footnote{Unfortunately we do not guarantee perfect emulation of earlier formats, for example if we implement bug fixes which affect how earlier versions are dealt with.}  
The version number is of the form \verb|<number>.<letter1>.<letter2>|.  The number corresponds to a major {\sc Essence} version number, and the first letter the {\sc Essence}$'$ revision within that.
Finally, the second letter indicates changes which are meant to be backwards compatible.

Parameter declaration, Constant definitions and Variable 
declarations can be interleaved, but for readability we suggest to put them in the 
order given above. Comments are preceded by `\$'.

Parameter values are defined in a separate file, the 
{\em parameter file}. Parameter files have the same header 
as problem models and hold a list of parameter definitions.
Table \ref{tab:modelstructure} gives an overview of the model
structure of problem and parameter files.
Each model part will be discussed in more detail in the following sections.

\begin{table}
\begin{center}
\begin{tabular}{|l||l|}
\hline
Problem Model Structure & Parameter File Structure \\
\hline
\hline
%\$ problem model               & \$ parameter file \\
{\tt language ESSENCE' 1.b.a}  & {\tt language ESSENCE' 1.b.a}  \\
\ & \\
\$ {\em parameter declaration}       & \$ {\em parameter instantiation} \\
{\tt {\bf given} } $n$ : int         & {\tt {\bf letting} } $n$ {\tt be} 7 \\
\$ {\em constant definition}        & \\
{\tt {\bf  letting}} $c$ {\tt be} 5   & \\
\ & \\
\$ {\em variable declaration }       & \\
{\tt {\bf find} } $x,y$ : int(1..$n$)  & \\
\ & \\
\$ {\em constraints}                 & \\
{\tt{\bf such that} } & \\
\ \ $x$ + $y$ $>=$ c,    & \\  
\ \ $x$ + $c$*$y$ $=$ 0    & \\  
\hline
\end{tabular}
\caption{Model Structure of problem files and parameter files in {\sc
    Essence}$'$. `\$' denote comments.}\label{tab:modelstructure}
\end{center}
\end{table}

\subsection{Constant Definitions}
In most problem models there are re-occurring constant values and
it can be useful to define them as constants. The {\tt letting}
statement allows to assign a name with a constant value. The statement
\begin{center}
{\tt letting } {\em NAME} {\tt be} {\em constant}
\end{center}
introduces a new reserved name {\em NAME} that is associated with 
the constant value {\em constant}. Every subsequent occurrence of 
{\em NAME} in the model is replaced by {\em constant}. Please note 
that {\em NAME} cannot be used in the model {\em before} it has been 
defined. In the following subsections we discuss different kinds of 
constants.

\subsubsection{Constant Expressions}
The statement
\begin{center}
{\tt letting } {\em c} {\tt be} 10
\end{center}
introduces a new constant with name $c$ that is assigned the value
10. Usually constants are written with lower-case letters. The
constant expression may also contain other constants or parameter
values, for instance with
\begin{center}
\begin{tabular}{llll}
{\tt letting } & {\em c}& {\tt be}& 10\\
{\tt letting } &{\em d} &{\tt be}& $c$*2\\
\end{tabular}
\end{center}

\subsubsection{Constant Domains}
Constant domains are defined in a similar way using 
the keywords {\tt be domain}:
\begin{center}
{\tt letting } {\em INDEX} {\tt be domain} $int$(1..5)
\end{center}
defines a domain with name {\em INDEX} that ranges 
from $int$(1..5).
For readability we suggest to use upper-case letters for domains.
 Constant domains may contain other 
constant/parameter values, such as in
\begin{center}
\begin{tabular}{llll}
{\tt letting } & {\em c}& {\tt be} &10\\
{\tt letting } &{\em INDEX}& {\tt be domain} &$int$(1..$c$) \\
\end{tabular}
\end{center}

\subsubsection{Constant Arrays}
Constant arrays are defined by stating a name, followed by 
the corresponding matrix type and the constant values in brackets.
For instance,
\begin{center}
{\tt letting }   $name_1$ :  {\tt matrix indexed by } [ $int$(1..4)] of $int$(1..10)  {\tt be}   [2,8,5,9]\\
\end{center}
defines a 1-dimensional constant array $name_1$ with 4 Elements
ranging from $int$(1..10). Hence  $name_1$[3] is replaced by 5.
2-dimensional constant arrays are defined in the same way:
\begin{center}
\begin{tabular}{ll}
{\tt letting }&  $name_2$ : \\
      &\ \ \ \ \ \  {\tt matrix indexed by } [ $int$(1..2), $int$(1..4)] of $int$(1..10) \\
 & \ \ \ \ \ \   {\tt be}  [ [2,8,5,1], \\
 & \ \ \ \ \ \ \ \ \ \ \ \ [3,7,9,4] ]   \\
\end{tabular}
\end{center}
states that $name_2$ is a (1..2)$\times$(1..3)array of integers ranging from
$int$(1..10) consisting of the values [ [2,8,5], [3,7,9] ]. Hence
$name_2$[1,2] is replaced by 8. 




\subsection{Parameter Declarations}
Parameters are declared with the {\tt given} statement followed 
by a domain the parameter ranges over. Parameters are allowed to 
range over the infinite domain $int$. As example, consider
\begin{center}
\begin{tabular}{lll}
{\tt given} & $n$ & : int \\
%{\tt given} & $D$ & : int(1..$n$) \\
\end{tabular}
\end{center}


\subsection{Variable Declaration}
Variables are declared using {\tt find} followed by a name and their
corresponding domain. The example below
\begin{center}
{\tt find}  $x$ : $int$(1..10)
\end{center}
defines a variable $x$ on the domain $int$(1..10).
It is possible to define several variables on the same domain as 
\begin{center}
{\tt find}  $x,y,z$ : $int$(1..10)
\end{center}
that introduces 3 variables $x,y,z$ ranging over $int$(1..10).
Arrays of variables are declared by defining the index domain and
basedomain. Consider the example
\begin{center}
{\tt find}  $m$ : {\tt matrix indexed by} [$int$(1..10)] of {\tt bool}
\end{center}
that declares $m$ as a 1-dimensional matrix of 10 Booleans. 


\subsection{Objective}
The objective of a problem is either to {\tt maximise} or 
{\tt minimise} a variable or expression. For instance,
\begin{center}
{\tt minimise} $x$
\end{center}
states that the value assigned to variable $x$ will be minimised.


\subsection{Constraints}
After defining constants and declaring variables and parameters,
constraints are specified with the keyword {\tt such that}.
 {\sc Essence}$'$ supports a wide range of operators (for more details
on operators see Section \ref{sec:op}):

\begin{itemize}

\item Basic Arithmetic Operators: $+$ \  $-$ \  * \  / \ \% \  $\mid$ {\tt
  min} \ {\tt max} 
\item Basic Boolean Operators: \verb1\1/ \  /\verb1\1 \  $=>$ \  $<=>$
\item Relational Operators: = \  != \  $>$ \  $<$ \  $>=$ \  $<=$
\item Sum Operator: {\tt sum}
\item Quantification operators: {\tt forall } \ {\tt exists}
\item Global Constraints: {\tt alldifferent} \ {\tt element}
\item Table Constraint: {\tt table}
\end{itemize}

We define two kinds of expressions: arithmetic and relational
expressions. Arithmetic relations range over an integer domain, 
for instance $x + 3$ is an arithmetic expression ranging from 
($lb(x)+3 .. ub(x)+3$). Relational expressions range over the 
Boolean domain, for instance the relational expression
 $x = 3$ can either be {\em true}
or {\em false}.
Basic arithmetic operators and the {\tt sum} operator produce 
arithmetic expressions and all other operators produce relational 
expressions.

Please note that each operator has a certain precedence and you will 
sometimes need to use parenthesis to express
certain constraints. As an example, consider the expression
\begin{center}
$x$ = $y$ /\verb1\1 $y$ $<$= $z$
\end{center}
According to the operator precedence in {\sc Essence}$'$, the
expression would be parsed as 
\begin{center}
$x$ = (  ($y$ /\verb1\1 $y$)  $<$= $z$ )
\end{center}
which might not be intended. Setting parenthesis 
is helpful to ensure a certain meaning, as illustrated below:
\begin{center}
($x$ =  $y$)  /\verb1\1 ($y$  $<$= $z$ )
\end{center}
More details about operator precedence is given in Section \ref{sec:op}.




\subsubsection{The {\tt sum} Operator}
The {\tt sum} operator corresponds to the mathematical 
$\sum$ and has the following syntax:
\begin{center}
{\tt sum} {\em quantified-variable(s)} : {\em domain} . {\em expression} 
\end{center}
For example, if we want to take the sum from 1 to 10 we write
\begin{center}
{\tt sum} $i$ : $int$(1..10) . $i$ 
\end{center}
which corresponds to 
\begin{center}
1 + 2 + 3 + 4 + 5 + 6 + 7 + 8 + 9 + 10
\end{center}
Several quantified variables may be defined for a particular
quantification domain and {\tt sum}s can also be nested:
\begin{center}
{\tt sum} $i,j$ : $int$(1..10) . \\
\ \ \ \ {\tt sum} $k$ : $int$(i..10) . \\
\ \ \ \ \ \ \ \ $x$[$i$,$j$] * $k$ 
\end{center}



\subsubsection{Universal and Existential Quantification}
Universal and existential quantification are powerful means to 
write down a series of constraints in a compact way. 
Quantifications have the same syntax as {\tt sum}, but with 
{\tt forall} and {\tt exists} as keywords:
\begin{center}
{\tt forall}  {\em quantified-variable(s)} : {\em domain} . {\em expression}
\end{center}
For instance, the universal quantification 
\begin{center}
{\tt forall}  $i$ : $int$(1..5) . \\
\ \ \ \ $x$[$i$] = $i$
\end{center}
corresponds to the conjunction 
\begin{center}
($x$[1] = 1) /\verb1\1 ($x$[2] = 2) /\verb1\1 $\dots$  /\verb1\1  ($x$[5] = 5)
\end{center}
which could also be written as 
\begin{center}
$x$[1] = 1, \\
$x$[2] = 2,\\
 $\dots$ \\ 
$x$[5] = 5
\end{center}
An example for existential quantification is 
\begin{center}
{\tt exists}  $i$ : $int$(1..5) . \\
\ \ \ \ $x$[$i$] = $i$
\end{center}
that corresponds to the disjunction
\begin{center}
($x$[1] = 1) \verb1\1/ ($x$[2] = 2) \verb1\1/ $\dots$  \verb1\1/  ($x$[5] = 5)
\end{center}
Quantifications can range over several quantified variables and can 
be arbitrarily nested, as demonstrated with the {\tt sum} operator.

% --------------------------------------------------------------

\subsection{Operators in {\sc Essence}$'$}\label{sec:op}

Arithmetic operators and the sum operator return arithmetic
expressions while all other expressions return relational expressions.
Arithmetic operators may only be used on arithmetic expressions
with the exception of addition and substraction (i.e. Booleans 
may be added/subtracted).\\

To use operators correctly, you have to understand the precedence and associativity of operators. 
Table \ref{tab:precedence} 
describes the precedence and associativity of the operators, arranged by decreasing order of precedence (the operators on top have highest precedence).



\begin{table}  
    \begin{center}
    \begin{tabular}{|l|lc|}
 
      \hline
       Operator(s) & Functionality & Associativity \\
       \hline
       \hline
       ,   & comma & Left \\   
       \hline
       :   & colon & Left \\
       \hline
       ( ) & left and right parenthesis & Left \\
       \hline
       \verb1[ ]1 & left and right brackets & Left \\
       \hline
       ! & not & Right \\  
       \hline
       \verb1/\1 & and & Left \\ 
       \hline
       \verb1\/1 & or & Left \\ 
       \hline
        {\tt =>} & if (implication)  & Left \\ 
	\hline
        {\tt <=>} & iff (logical equality) & Left \\ 
	\hline
        {\tt -} & unary minus & Right \\
	\hline
        \verb1^1 & power & Left \\ 
	\hline
        * / & multiplication, integer division & Left \\ 
	\hline
        {\tt + -} & addition, substraction &  Left \\
	\hline
        {\tt < <= > >= } & (lex)less, (lex)less or equal, &  \\ 
	{\tt <lex <=lex >lex >=lex } & (lex)greater, (lex)greater or equal & none \\ 
	\hline
        {\tt = !=} & equality, disequality & none \\
	\hline
        . & dot & Right \\ 
	\hline

      \end{tabular}
\end{center}
    \caption{Operator precedence in {\sc Essence}$'$}

\label{tab:precedence}
\end{table}

As you would expect, operators with higher precedence take priority, so are applied first.   We have, for example 
\begin{eqnarray*}
a => b \verb.^. c * d / ( e + f ) & \equiv & a => ( (b \verb.^. c) * d ) /  (e + f) )
\end{eqnarray*}

For associativity, an operator $\cdot$ with left associativity has 
$a \cdot b \cdot c \equiv  a \cdot b) \cdot c$ while with right associativity we have 
$a \cdot b \cdot  c \equiv a \cdot (b \cdot c)$.   The operators with no associativity defined are meaningless if nested, so that $a = b = c$ is incorrect.    



\section{Examples} 

\subsection{EightNumberPuzzle.eprime}
\begin{verbatim}
language ESSENCE' 1.b.a

find circles: matrix indexed by [int(1..8)] of int(1..8)

such that

alldiff(circles),
| circles[1] - circles[2] | > 1,
| circles[1] - circles[3] | > 1,
| circles[1] - circles[4] | > 1,
| circles[2] - circles[3] | > 1,
| circles[3] - circles[4] | > 1,
| circles[2] - circles[5] | > 1,
| circles[2] - circles[6] | > 1,
| circles[3] - circles[5] | > 1,
| circles[3] - circles[6] | > 1,
| circles[3] - circles[7] | > 1,
| circles[4] - circles[6] | > 1,
| circles[4] - circles[7] | > 1,
| circles[5] - circles[6] | > 1,
| circles[6] - circles[7] | > 1,
| circles[5] - circles[8] | > 1,
| circles[6] - circles[8] | > 1,
| circles[7] - circles[8] | > 1
\end{verbatim}

\subsection{FarmersProblem.eprime}
\begin{verbatim}
language ESSENCE' 1.b.a

find pigs, hens: int(0..7)

such that

pigs + hens = 7,
pigs * 4 + hens * 2 = 22
\subsection{K4P2GracefulGraph.eprime}
\begin{verbatim}
language ESSENCE' 1.b.a

find nodes : matrix indexed by [int(1..8)] of int(0..16),
       edges: matrix indexed by [int(1..16)] of int(1..16)

such that

|nodes[1] - nodes[2]| = edges[1],
|nodes[1] - nodes[3]| = edges[2],
|nodes[1] - nodes[4]| = edges[3],
|nodes[2] - nodes[3]| = edges[4],
|nodes[2] - nodes[4]| = edges[5],
|nodes[3] - nodes[4]| = edges[6],

|nodes[5] - nodes[6]| = edges[7],
|nodes[5] - nodes[7]| = edges[8],
|nodes[5] - nodes[8]| = edges[9],
|nodes[6] - nodes[7]| = edges[10],
|nodes[6] - nodes[8]| = edges[11],
|nodes[7] - nodes[8]| = edges[12],

|nodes[1] - nodes[5]| = edges[13],
|nodes[2] - nodes[6]| = edges[14],
|nodes[3] - nodes[7]| = edges[15],
|nodes[4] - nodes[8]| = edges[16],

alldiff(edges),
alldiff(nodes)
\end{verbatim}
\subsection{NQueensBool.eprime}
\begin{verbatim}
language ESSENCE' 1.b.a

given n: int
find queens: matrix indexed by [int(1..n), int(1..n)] of bool

such that

forall i : int(1..n). ( sum j : int(1..n). queens[i,j] ) = 1,
forall i : int(1..n). ( sum j : int(1..n). queens[j,i] ) = 1 ,

forall i : int(1..n). forall j: int(1..n). forall k:int(1..n). forall l:int(1..n).
(
  ( 
    (|i - k| = |j - l| ) /\ 
    ((i != k) \/ (j != l)) 
  ) 

=> 

   (queens[i,j] + queens[k,l] <= 1)
)
\end{verbatim}
\subsection{NQueensColumn.eprime}
\begin{verbatim}
language ESSENCE' 1.b.a

given n: int
find queens: matrix indexed by [int(1..n)] of int(1..n)

such that

alldiff(queens),

forall i : int(1..n). forall j : int(i+1..n). | queens[i] - queens[j] | != |i - j|
\end{verbatim}
\subsection{SENDMOREMONEY.eprime}
\begin{verbatim}
language ESSENCE' 1.b.a


find S,E,N,D,M,O,R,Y : int(0..9)

such that

1000*S + 100*E + 10*N + D + 
1000*M + 100*O + 10*R + E =
10000*M + 1000*O + 100*N + 10*E + Y,

alldiff([S,E,N,D,M,O,R,Y])
\end{verbatim}
\subsection{zebra.eprime}
\begin{verbatim}
language ESSENCE' 1.b.a

$red = colour[1]
$green = colour[2]
$ivory = colour[3]
$yellow = colour[4]
$blue = colour[5]
$Englishman = nationality[1]
$Spaniard = nationality[2]
$Ukranian = nationality[3]
$Norwegian = nationality[4]
$Japanese = nationality[5]
$coffee = drink[1]
$tea = drink[2]
$milk = drink[3]
$orange juice = drink[4]
$Old Gold = smoke[1]
$Kools = smoke[2]
$Chesterfields = smoke[3]
$Lucky Strike = smoke[4]
$Parliaments = smoke[5]
$dog = pets[1]
$snails = pets[2]
$fox = pets[3]
$horse = pets[4]


find colour: matrix indexed by [int(1..5)] of int(1..5),
       nationality: matrix indexed by [int(1..5)] of int(1..5),
       drink: matrix indexed by [int(1..5)] of int(1..5),
       smoke: matrix indexed by [int(1..5)] of int(1..5),
       pets: matrix indexed by [int(1..5)] of int(1..5)

such that

$constraints needed as this is a logical problem where 
$the value allocated to each position of the matrix represents positon of house
alldiff(colour),
alldiff(nationality),
alldiff(drink),
alldiff(smoke),
alldiff(pets),

$There are five houses.
$No constraint covered by domain specification

$The Englishman lives in the red house
nationality[1] = colour[1],

$The Spaniard owns the dog.
nationality[2] = pets[1],

$Coffee is drunk in the green house.
drink[1] = colour[2],

$The Ukranian drinks tea.
nationality[3] = drink[2],

$The green house is immediately to the right of the ivory house.
colour[2] + 1 = colour[3],

$The Old Gold smoker owns snails.
smoke[1] = pets[2],

$Kools are smoked in the yellow house.
smoke[2] = colour[4],

$Milk is drunk in the middle house.
drink[3] = 3,

$The Norwegian lives in the first house
nationality[4] = 1,

$The man who smokes Chesterfields lives in the house next to the man with the fox.
|smoke[3] - pets[3]| = 1,

$Kools are smoked in the house next to the house where the horse is kept.
|smoke[2] - pets[4]| = 1,

$The Lucky Strike smoker drinks orange juice.
smoke[4] = drink[4],

$The Japanese smokes Parliaments.
nationality[5] = smoke[5],

$The Norwegian lives next to the blue house.
|nationality[4] - colour[5]| = 1
\end{verbatim}
\subsection{zebraAlt.eprime}

\begin{verbatim}
language ESSENCE' 1.b.a

$red = colour[1]
$green = colour[2]
$ivory = colour[3]
$yellow = colour[4]
$blue = colour[5]
$Englishman = nationality[1]
$Spaniard = nationality[2]
$Ukranian = nationality[3]
$Norwegian = nationality[4]
$Japanese = nationality[5]
$coffee = drink[1]
$tea = drink[2]
$milk = drink[3]
$orange juice = drink[4]
$Old Gold = smoke[1]
$Kools = smoke[2]
$Chesterfields = smoke[3]
$Lucky Strike = smoke[4]
$Parliaments = smoke[5]
$dog = pets[1]
$snails = pets[2]
$fox = pets[3]
$horse = pets[4]


find colour: matrix indexed by [int(1..5)] of int(1..5),
       nationality: matrix indexed by [int(1..5)] of int(1..5),
       drink: matrix indexed by [int(1..5)] of int(1..5),
       smoke: matrix indexed by [int(1..5)] of int(1..5),
       pets: matrix indexed by [int(1..5)] of int(1..5)

such that

$constraints needed as this is a logical problem where 
$the value allocated to each position of the matrix represents positon of house
alldiff(colour),
alldiff(nationality),
alldiff(drink),
alldiff(smoke),
alldiff(pets),

$There are five houses.
$No constraint covered by domain specification

$The Englishman lives in the red house
nationality[1] = colour[1],

$The Spaniard owns the dog.
nationality[2] = pets[1],

$Coffee is drunk in the green house.
drink[1] = colour[2],

$The Ukranian drinks tea.
nationality[3] = drink[2],

$The green house is immediately to the right of the ivory house.
colour[2] + 1 = colour[3],

$The Old Gold smoker owns snails.
smoke[1] = pets[2],

$Kools are smoked in the yellow house.
smoke[2] = colour[4],

$Milk is drunk in the middle house.
drink[3] = 3,

$The Norwegian lives in the first house
nationality[4] = 1,

$The man who smokes Chesterfields lives in the house next to the man with the fox.
(smoke[3] = pets[3]+1) \/ (smoke[3] = pets[3]-1),

$Kools are smoked in the house next to the house where the horse is kept.
(smoke[2] = pets[4]+1) \/ (smoke[2] = pets[4]-1),

$The Lucky Strike smoker drinks orange juice.
smoke[4] = drink[4],

$The Japanese smokes Parliaments.
nationality[5] = smoke[5],

$The Norwegian lives next to the blue house.
(nationality[4] = colour[5]+1) \/ (nationality[4] = colour[5]-1)
\end{verbatim}

\begin{thebibliography}{10}

\bibitem{java} 
  Java 
 \newblock  http://java.sun.com

\end{thebibliography}

\end{document}


